A multi-scale story of the diffusion of a new technology: the web

### 2025-03-12 paper/submission2/response.qmd

---
title: "Response to editor and reviewer comments"
format: pdf
bibliography: bibliography.bib
---

Editor and reviewer comments as well as author responses are presented below. 
Comments are in \color{teal}teal fonts \color{black} and responses in black.

\color{teal}
Thank you for submitting your manuscript to Journal of Economic Geography.

@@

I have now received comments from reviewers. Their comments are at the end of this email. The reviewers raise a range of concerns and have questions about the suitability of the paper for publication. The overall feeling is that the paper has probably been submitted too early and needs major work on most sections if it is to become publishable. I concur with the reviewers and see a number of significant issues that need to be addressed. This left me with a dilemma in terms of the decision. I have decided to invite you to revise and resubmit the paper so that it can be considered further for publication. However, this is a very high-risk revision, and I was bordering on rejection given the scale of the work needed and the limitations that the reviewers highlight. I am saying this to be open about the risk that after the next round the paper could be rejected unless reviewers believe you have addressed the range of issues identified. You may want to factor this into your decision about whether to revise the paper, given how much work is needed.

Having said this, let me offer some suggestions about the main areas to focus on if you decide to revise the paper. Reviewer 1 is more positive but very brief in their report. I am, therefore, putting more weight on reviewers 2 and 3 when identifying priorities for the revision.
\color{black}

Thank you for providing the opportunity to revise and resubmit my paper. The 
reviewer reports were really helpful and I think it is fair to say that the
reviewers did go the extra mile to offer useful recommendations. I think I have 
adequately addressed their concerns and my feeling is that the paper has now 
been drastically improved. I hope the editor and the reviewers will agree.

\color{teal}
First, you will need to develop a more convincing theoretical argument and contribution. You will see that both reviewers 2 and 3 raise questions about your claims in relation to the gap in the literature and suggest you fail to build on existing work appropriately and do not convincingly identify the motivation for this analysis. In general, claims that an issue has never been addressed are hard to sustain and it is better to build on existing work and extend knowledge relating to a particular theoretical issue. The reviewers challenge you to do this using your data and suggest both the argument and structure of the paper will need to change.
\color{black}

The paper has been restructured to better illustrate the theoretical
argumentation, motivation and contribution. It is also better situated within 
the relevant literatures. Below I offer detailed responses to the reviewer 
comments.

\color{teal}
Second, you need to justify your data and analysis better. Reviewer 2 fairly questions the end point of the analysis – which is not explained and given the 10+ years gap to your submission raises questions about the timeliness of the analysis. The questions about coverage of the UK website population also need to be addressed convincingly. More generally, the reviewers ask for more careful presentation of the data to make clear the significance of each figure and finding. As it stands, the paper feels rather overloaded with data, rather than acting as a focused analysis.
\color{black}

The data issues have been clarified.

\color{teal}
The problems identified above result in the paper feeling imprecise, confusing and lacking a contribution. The end point is a conclusion that the reviewers do not find justified or original. The reviewers do see potential. As I noted, the feeling is a paper that is underdeveloped and submitted prematurely. There is, then, a route to publication, but it will require significant changes to convince the reviewers.
\color{black}

The restructuring of the paper around two hypotheses as well as the emphasis on
the granularity of the analysis and the the long time series approach allowed to
more clearly highlight the contribution in the conclusion section.

As a sidenote, there was a minor error in the previous version of the paper 
with the implementation of the spatially sensitive CV. This has now been 
corrected hence the minor changes in the modelling results (second decimal).

@@

## Referee: 1

\color{teal}
I have read your paper with interest and I appreciate its quality and clarity. The work is both well-executed and well-written, making it a valuable contribution to the field.

However, I would like to offer a few minor suggestions that could further enhance the readability and overall coherence of the paper:
\color{black}

Thank you for your nice comments and suggestions.

\color{teal}
Simplify the Introduction: I recommend streamlining the introduction by reducing the inclusion of overly technical or highly specific details. Instead, consider expanding on the broader conceptual framework that situates the paper within its scholarly context. This approach would help orient readers and provide a more accessible foundation for understanding the study.
\color{black}

Both the introduction and the literature review sections have been restructured
and simplified. Some of the concepts initial unpacked in the introduction
have now moved to the second section.

\color{teal}
Strengthen the Connection Between Hypotheses and Results: It may be beneficial to establish a clearer and more explicit link between the study's hypotheses and its findings. One way to achieve this could be by using section titles that are more conceptually informative, which would guide the reader through the logical progression of the argument and findings more effectively.
\color{black}

The second section has been restructured and is now based on two hypotheses, 
which are also used to structure the methodology and the results sections. All 
section titles have been changed:

- The mechanisms that shape the diffusion of a new technology
- Results
  - The three diffusion mechanisms at granular geographical scales
  - Early adoption and latecomers
- Conclusions

@@

## Referee: 2

\color{teal}
OVERVIEW\
- focused on innovation diffusion, using the case of the web in the UK
- “this paper tests whether the three distinct spatial diffusion mechanisms shaped the diffusion of the Web in the UK at very granular spatial scales from the Web’s early stages onwards”; presumably “S-shaped pattern”/temporal models, core to periphery; distance/proximity.

Figure 2 shows the diffusion of domains per company at the scale of the UK matches the S growth model. Figure 3 shows some spatial clustering. I think this is for all years in the dataset but this could be better labeled
\color{black}

Both figures has been re-titled:

- Cumulative adoption of the Web, UK
- LAD Web diffusion speed, 1996-2012

Please note that the old Figure 3 has now moved to the end of the Results 
section (now Figure 8).

@@

\color{teal}
Figure 4, not sure what this demonstrates
\color{black}

This figure has also moved towards the end of the Results section as it is now 
used to test the second hypothesis. The new title reads "Figure 9: Dynamics of 
Wed diffusion based on LAD ranking" and the following paragraph explains how it 
was created and what it reveals.

> Next, the stability and volatility of these patterns is analysed to reveal 
patterns of early adoption and leapfrogging. Figure 
9 plots the relative ranking of LAD based on website density at the
beginning and the end of the study period. To decrease noise, the average 
ranking of 1996-1998 and 2010-2012 is plotted instead of the individual years. 
Each line represents a single LAD and the colours depict the 
5\textsuperscript{th} and 95\textsuperscript{th} percentile of the absolute
difference of the LAD ranking between 1996-1998 and 2010-2012. In other words, 
the 5\textsuperscript{th} percentile includes the LAD, the relative
position of which remained unchanged between the beginning and the end of 
the study period, both at the top and at the bottom of the hierarchy (small
absolute difference in ranking). The 95th percentile includes LAD that either
increased or decreased a lot their relative positions over time (high absolute 
difference in ranking). Thus, Figure 9 reveals cases of 
*leapfrogging* since the LAD at the top-right part of the figure managed to 
jump at the top of the hierarchy at the end of the study period.

@@

\color{teal}
Figure 5&6, shows early clustering and then diffusion (I think)
\color{black}

Yes. This is also reflected in the text. For instance:

> This reflects the early concentration of high website density around London, which over time diffused as high website density clusters can be seen in other parts of the country away from London (Figure 5)

> Eventually, as the adoption rate increased, more such clusters of high website density were formed and this is reflected both in the Moran’s I and the LISA maps in Figures 4 and 5. A similar pattern is observed for the extended dataset (see FiguresA.4 and A.5).

@@

\color{teal}
DATA\
Focusing just on .co.uk website is OK but needs to be better justified given that it only represented about 50 percent of domains in the UK at that time. See my back of the envelop calculations: As per Zook (2001, Table 1) .uk represented 66% of all domains in the UK, .co.uk is a subset of that, so this is perhaps 50% of all domains registered in the UK at that time. If 1/5 of .co.uk domains had findable postal codes (p. 5), this study is based on 10% of the domains in the UK. (perhaps this is too low, aggregating from webpages to websites brought the number higher, not entirely clear to me)
\color{black}

Thanks for going the extra mile to check the data used here. I agree with the 
above calculation up until the findable postcode point as 1/5 of the archived 
*webpages* -- not *websites*, i.e. domains -- included a postcode in the web text. 
Think of the website of a small company with one trading address. I think it is 
fair to expect a website to have multiple webpages to present the different
products and services and then one webpage to showcase the address -- the 
'Contact us' page. So, the 1/5 share of *webpages* containing a postcode should
translate to a much higher share of *websites* -- i.e. domains -- containing 
a postcode. Previous estimates indicated that around 70% of all websites contain 
place reference [@hill2009georeferencing] and this reference has now been added 
in the paper.

So, based on the above 35% of all domains with a registered address in the UK
were represented in the data. Most importantly though the paper does not model 
the adoption of the Web from the general population, but instead from firms as 
only commercial websites are considered. So, although the data only represents 
35% of the population of websites with a registered address in the UK, it
represents almost all the *relevant* archived websites -- that is websites using
the commercial .co.uk SLD. As stated in the paper, this has been the primary 
choice for commercial websites in the UK.

@@

> Instead of using the whole .uk ccTLD, this paper focuses on its
commercial subset, the .co.uk second level domain (SLD). This choice
is aligned with the topic of the paper -- the diffusion of the Web for 
commercial functions and decreases the heterogeneity of the web data as such 
commercial websites have specific aims: they are used to diffuse information, 
support online transactions and share opinions [@THELWALL2000441; 
@blazquez2018big].

<!-- Active/inactive, don;t suffer from multiple registrations -->

\color{teal}
overall the time period 1996-2012 is a good time to study the diffusion of the commercial internet. But why stop at 2012? Is it when the Library stopped its data series?
\color{black}

Yes, the British Library had obtained a copy of all of the .uk contents of the 
Internet Archive for that period. From a data perspective, it wouldn't be 
correct to attempt to extend this dataset with data from other open web archives,
such as the Common Crawl, as they use different crawlers and there will be
inconsistencies. Most importantly though, this period covers the diffusion of 
the web from the very early stages until its maturity. This is also illustrated 
by the figures depicting the cumulative adoption and the flattening right end of 
the curve for both subsets of the data (Figure 2 and A1).

There is a specific paragraph justifying how the time period of the data fits
the aim of the paper and another line is now included in the second paragraph 
of the introduction ("from the very early days until its maturity (1996-2012)").

@@

\color{teal}
Framing
“With the exception of Sinai and Waldfogel (2004), who focused, among other things, on web content and demonstrated that more such content is available in larger markets, this strand of literature has been mostly concerned with the spatiality of the Internet’s hard and tangible infrastructure and shied away from exploring the spatial dimensions of one of the most distinct software layers of the Internet infrastructure, the Web.” This statement is simply not true, the literature reviews misses the entirety of web content diffusion studies: Zook, 2000, Malecki, 2002. Townsend, 2000, Gorman, 2002, Feldman, 2002, lots more
\color{black}

Point taken. Despite being well aware of these studies, I omitted citing them as 
I was primarily searching for studies testing the diffusion mechanisms over 
longer time periods. This error has now been corrected and the above paragraph 
has been replaced with the following one, which emphasises the long time series 
motivation. Regarding the actual references, @zook2000web, @zook2001old and 
@moss1997tracking used relevant, domain name registration data, but the rest 
(e.g. @malecki2002economic and @moss2000internet) focused on hard 
infrastructure. @gorman2002web examined the head and branch locations of the top 
40 e-business integration firms in the USA, which although interesting is not 
directly related to this paper. I also went through Feldman's work from around 
2002, but I did not find papers adopting a similar empirical perspective.

> Another line of motivation has to do with the very nature of the Web as a 
digital technology and how it evolved over time. The digital geographies field 
has debunked premature claims about the Internet's placeless-ness by mapping 
various of its facets and technological layers [@tranos2013death], but the lack 
of relevant long time-series data prevented researchers from observing digital
phenomena overtime. Most of these studies focused on its hard infrastructure 
[e.g. @malecki2002economic; @moss2000internet], on Internet subscribers 
[e.g. @blank2018local], or on social media users [e.g. @crampton2013beyond].
Exceptions include early work from @zook2000web, @zook2001old and 
@moss1997tracking, which utilised domain name registration data and use it to
map and illustrate the spatial footprint of the web content. Pivotal as these 
contributions have been in understanding the economic geographies of the 
Internet, they were based on cross-sectional, often spatially detailed, mapping 
and on short temporal comparisons (e.g. 1993-1998) and did not observe the 
evolution and diffusion of the Web until its maturity. As we know from the 
relevant literature, different technologies diffuse at different rates and 
there is no theory that can explain these differences 
[@leibowicz2016representing]. Hence, mapping and modelling the diffusion of the
Web can offer useful insights.

@@

\color{teal}
“for a digital technology as weightless as the Web” – this runs a bit counter to lots of work by digital geographers who have shown again and again how the digital is not weightless or placeless
\color{black}

See response to the next comment.

\color{teal}
Conclusion
p. 18 “This is the first time that the importance of granular spatial mechanisms for the diffusion of an intangible, digital technology is exposed.” Honestly, I would not expect the diffusion of the web to be any different than any other kind of innovation diffusion. The spatiality of digital life, and its connection to material spaces is a constant refrain within digital geography. So I’m not convinced it makes much sense to emphasize this, if a granular spatial analysis of innovation diffusion is already novel.
\color{black}

Point taken. The emphasis now is on the granularity of the analysis and also on 
the long time series that allow to observe the technology almost from its 
beginning until its maturity. The above text as well as other similar text 
elsewhere in the paper have been amended accordingly.

@@

## Referee: 3

\color{teal}
This paper addresses a gap in Economic Geography by examining the spatial diffusion of digital technologies, specifically the adoption of commercial websites, from 1996 to 2012. Recognizing the potential of digital technologies to shape local development trajectories through increased productivity and innovation, the study employs novel data sources and machine learning techniques to analyze diffusion patterns at highly granular geographical scales. The authors explore the roles of various spatial mechanisms driving this process, offering insights into how digital technology adoption unfolds across space and time.
The paper has the potential to make an interesting contribution to the literature as it combines a fascinating topic with novel data and methodologies. However, in its current form, it is not ready for publication, as it never becomes clear what it actually delivers in terms of new knowledge and what knowledge gap it seeks to fill. Despite some analytical elements, it is rather descriptive in nature, which doesn’t fit well with this journal.

Most importantly, I suggest developing some clear hypotheses to structure and motivate the paper better. In light of this, I recommend major revisions.
\color{black}

This is very helpful, thank you. The whole paper has now been restructured
around two hypotheses. I believe the structure and the flow of the paper are 
much clearer now.

@@

\color{teal}
\noindent 1. Abstract:\
\noindent 1. The abstract needs significant improvement. It needs to adequately report the main findings, relevance, or implications of the study. Additionally, the assertion that geographers have "shied away" from analyzing the diffusion of digital technologies is misleading; the issue was primarily due to the lack of available data.
\color{black}

The abstract has been amended accordingly:

> This paper investigates the spatial diffusion of a new technology that is the Web in the UK. It employs novel data and machine learning methods to model the influence of well-established diffusion mechanisms. Contrary to previous studies, it adopts multiple scales, high spatial granularity and a long study period that captures the early stages of the Web until its maturity (1996-2012). Findings reveal the importance of such spatial mechanisms (namely distance, urban hierarchy and the S-shaped pattern of the cumulative level of adoption) even at granular scales. They also highlight spatial heterogeneity and instances of leapfrogging.

@@

\color{teal}
\noindent 2. Introduction\
\noindent 2. General Diffusion Studies: The statement that there are few contemporary general diffusion studies in Economic Geography is not accurate, although it may apply specifically to digital technologies. Consider the following recent examples:
o Peris, A., Meijers, E., & van Ham, M. (2021). Information diffusion between Dutch cities: Revisiting Zipf and Pred using a computational social science approach. Computers, Environment and Urban Systems, 85(November), 101565. https://doi.org/10.1016/j.compenvurbsys.2020.101565
o Bednarz, M., & Broekel, T. (2020). Pulled or pushed? The spatial diffusion of wind energy between local demand and supply. Industrial and Corporate Change, 29(4), 893–916. https://doi.org/10.1093/icc/dtaa012
o Feldman, M. P., Kogler, D. F., & Rigby, D. L. (2015). rKnowledge: The Spatial Diffusion and Adoption of rDNA Methods. Regional Studies, 49(5), 798–817. https://doi.org/10.1080/00343404.2014.980799
\color{black}

@PERIS2021101565 is indeed an interesting and, to humble opinion, novel paper, 
but it is not directly relevant as it focuses on information diffusion rather 
than diffusion of new technologies. Information diffuses through social and, 
consequently, urban networks, but it is not adopted like new technologies. 
Then, @bednarz2020pulled was already cited and @feldman2015rknowledge has now 
been added. These references are included in the second section.

Regarding the statement about economic geography studies on diffusion of 
new technologies and innovation, respectfully, I do believe that this is a fair
statement. The references listed below are the ones reviewed in the second 
section and none of them are published in one of the flagship economic geography
journals (*Journal of Economic Geography* and *Economic Geography*). There are 
two papers published in *The Professional Geographer*, which is a broad-scope
geography journal and then only one paper in *Regional Studies*, which is the 
closest to economic geography. Importantly though, the *Regional Studies* paper 
[@feldman2015rknowledge] is focused on inventions as observed by patent 
data contrary to the broader 'adoption of a new technology' approach of this 
paper. All other papers listed below are published in business and economics 
journals.

@@

1. @beardsell1999spatial: *European Economic Review*
2. @bednarz2020pulled: *Industrial and Corporate Change*
3. @haller2011determinants: *Applied Economics*
4. @perkins2011internet: *The Professional Geographer*
5. @feldman2015rknowledge: *Regional Studies*
6. @ding2010modeling: *The Professional Geographer*
7. @bakher2013diffusion: *Journal of Marketing Analytics*
8. @PAPAGIANNIDIS2015308: *Technological Forecasting and Social Change*
9. @lengyel2020role: *Scientific reports*

Also, the lack of spatial granularity argument has been highlighted in various 
instances throughout the paper as the main line of justification.

\color{teal}
\noindent 3. Novelty: The study needs to be more explicit about its novel contributions to understanding spatial diffusion mechanisms. Currently, the discussion is too general.
\color{black}

The novelty of the paper lies in the fact that there are hardly any studies
that analysed the role of these three spatial mechanisms in the diffusion 
of a new technology at such granular spatial scales. Importantly, the paper does 
this by adopting a long time series perspective which captures the commercial 
beginning of the Web until its maturity. The following paragraph in the 
introduction highlights this:

### 2025-03-12 paper/submission2/anonymised.qmd
paper v2

---
title: "A multi-scale story of the diffusion of a new technology: the web"
author:
  - name: Emmanouil Tranos
    email: e.tranos@bristol.ac.uk
    affiliations: 
      - id: some-tech
        name: University of Bristol
        department: 
        address: 
        city: 
        state: State
        postal-code: UK
    corresponding: e.tranos@bristol.ac.uk
address:
  - attributes:
        corresponding: true
    note: This is the first author footnote.
abstract: |
    Despite the broad consensus that differences in technological adoption and sophistication can lead to uneven trajectories of local development as new technologies can increase productivity and allow firms to introduce new products and services, economic geographers have shied away from exploring the diffusion of digital technologies across space and time. This paper fills this gap by exposing the roles of different spatial mechanisms in the diffusion of an intangible, digital technology at very granular geographical scales. It does so by employing a novel source of data and machine learning algorithms to model the active engagement with Web during 1996-2012.
keywords: 
  - diffusion
  - technologies
  - Web
  - digital
date: last-modified
bibliography: bibliography
format:
  elsevier-pdf:
    keep-tex: true
    text: |
      \usepackage[demo]{graphicx} % "demo" option just for this example
      \usepackage{subcaption}
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: authoryear
editor: 
  markdown: 
    wrap: 72
---

@@

<!-- ADD SOMETHING ABOUT MULTIPLE SCALES -->

<!-- geographers lack of interest for diffusion + motivation -->
Economic geographers were always interested in how new technologies and
innovations diffuse across space and time with the work of
@hagerstrand1968innovation, which demonstrated the importance of
communications as a mechanism for the diffusion of innovations, being
the most impactful. It was followed by various empirical studies -- see
for instance @ormrod1990 and @iso2005. However, the torch of exploring
and modelling such diffusion processes had been passed to other
disciplines such as economics, business studies and sociology well
before the 'cultural turn' of economic geography
[@perkins2005international; @ding2010modeling].
<!-- NEW spatial granularity --> As a consequence, relevant studies
did not investigate diffusion mechanisms at granular geographical scales to 
provide local policy insights, for instance, regarding infrastructural d
evelopment and local technology policies. 
A potential explanation of the lack of geographical studies exploring the 
diffusion of new and, more specifically for this paper, digital technologies 
across both space and time can be attributed to the scarcity of relevant and 
granular enough data, a problem also highlighted by @iso2005 and
@kemeny2011international. As @zook2022mapping pointed, digital
activities are hardly ever captured in official data.

<!-- contribution -->
This paper offers such a contribution: an economic geography study
revealing the importance of spatial mechanisms for the diffusion of an
intangible, digital technology that is the Web in the UK at a high level
of spatial granularity 
<!-- NEW TO JUSTIFY DATA TIME PERIOD-->
from the very early days until its maturity (1996-2012).
<!-- NEW DATA CLARIFICATION -->
It specifically focuses on the diffusion of commercial websites.
It does so by employing a novel source of big data and machine learning
algorithms to model the *active* engagement with Web -- that is creating and 
maintaining instead of just browsing a website -- during that period
and tests how different spatial and temporal mechanisms shaped its
diffusion. This paper also exemplifies how the combination of data
sources and state-of-the-art research methods which escape the
traditional social science domain can offer new lenses to geographical
research regarding the understanding of technological diffusion.

<!-- motivation: stakeholders -->
The motivation for this paper lies in the fact that there are various
stakeholders who are interested in knowing how new digital technologies
diffuse over space and time and use this knowledge to make predictions
about the diffusion of related future technologies. As per
@leibowicz2016representing, historical studies agree that technologies
diffuse differently in terms of times, rates, and geographies and can be
driven by related policies [@victor1993]. @meade2021modelling highlight
that a variety of actors have a direct interest in gaining such
knowledge including network equipment suppliers, network operators,
regulatory and local authorities. These processes and their effects vary
a lot across scales: although the diffusion of a new technology might
not be optimal at a local level, it might be beneficial from a global
perspective as it could lead to faster diffusion to less advantaged
places [@leibowicz2016representing]. Despite the spatial heterogeneity
of such diffusion mechanisms and the policy relevance, there is a
scarcity of studies analysing the diffusion of new digital technologies
at a detailed geographical level because of the lack of detailed enough
data and adequate research methods.

@@

<!-- NEW motivation: intangible/digital -->
Another line of motivation has to do with the very nature of the Web as a 
digital technology and how it evolved over time. The digital geographies field 
has debunked premature claims about the Internet's placeless-ness by mapping 
various of its facets and technological layers [@tranos2013death], but the lack 
of relevant long time-series data prevented researchers from observing digital
phenomena overtime. Most of these studies focused on its hard infrastructure 
[e.g. @malecki2002economic; @moss2000internet], on Internet subscribers 
[e.g. @blank2018local], or on social media users [e.g. @crampton2013beyond].
Exceptions include early work from @zook2000web, @zook2000web and 
@moss1997tracking, which utilised domain name registration data and use it to
map and illustrate the spatial footprint of the web content. Pivotal as these 
contributions have been in understanding the economic geographies of the 
Internet, they were based on cross-sectional, often spatially detailed, mapping 
and on shallow time comparisons (e.g. 1993-1998) and did not observe the 
evolution and diffusion of the Web until its maturity. As we know from the 
relevant literature, different technologies diffuse at different rates and 
there is no theory that can explain these differences 
[@leibowicz2016representing]. Hence, mapping and modelling the diffusion of the
Web among firms can offer useful insights.

@@

<!-- description of the methods and results -->
<!-- AMMEND I AFTER CHANGE THE RESULTS -->
To analyse these data, a diverse set of methods are employed. Firstly,
cumulative adoption curves are estimated for the UK as a whole and for
the Local Authority Districts (LAD), illustrating, for the first time,
such S-shaped diffusion patterns for local areas. This system level
analysis also reveals insights about the volatility of the adoption of
the Web. Then, exploratory spatial data analysis offers some first
indications of the importance of the different spatial diffusion
mechanisms in shaping the diffusion of the Web in the UK. Importantly,
the analysis adopts two distinct geographical scales: the LAD and the
most detailed scale of UK census areas, the Output Areas (OA). Lastly, a
novel methodological framework based on Machine Learning (ML) algorithms
is developed to test the role of such mechanisms at these two scales.
The results reveal, for the first time, the importance of different
spatial diffusion mechanisms in the diffusion of an intangible, digital
technology at very granular geographical scales.

<!-- contents -->
<!-- AMMEND I AFTER CHANGE THE RESULTS -->
The structure of the paper goes as follows. @sec-litreview reviews the
key concepts in the diffusion literature as well as relevant empirical
studies. Then, @sec-datamethods describes the data and methods used in
this paper. The results of the different stages of the analysis are
presented in @sec-results. The discussion of the results and the
conclusions of the paper are offered in @sec-conclusions.

@@

<!-- hierarchy -->
@hagerstrand1968innovation also emphasised the role of the top-down
hierarchy of urban systems and highlighted that new technologies and
innovations are firstly adopted in larger cities and then diffuse to
second tier ones. This is a sequential instead of a simultaneous
process, which resembles the 'lead-lag' spatial acceleration effect in
market research [@bento2018time; @PERES201091]. Schmidt's Law
empirically illustrates a similar pattern [@grubler1990rise]. Core and
highly agglomerated regions is where new technologies are invented and
commercially deployed. This is where the first adopters tend to be
based. Then, technologies spread to the rim and eventually to the
periphery. Although adoption pace might be higher when new technologies
finally arrive to the periphery, the saturation levels there may never
reach the ones in the core because of the lack of infrastructure or
other necessary institutions [@leibowicz2016representing].

@@

> H~2~: The diffusion of the commercial Web is shaped by both early adoption 
and late comer effects.

These are the two hypotheses are empirically tested in @sec-results using the 
data and methods described in the next section.

@@

[^2]: [https://archive.org/](See%20https://archive.org/).

@@

The analysis is firstly conducted on websites, which only contain one
unique postcode. As a robustness check, the analysis is replicated for
an extended subset of websites, which include up to 10 unique postcodes
to capture commercial websites which point to multiple locations. They
are geolocated by equally attaching them to these locations. This
extended sample includes 94 percent of all the archived websites in
2000.

@@

f.websites.pc <- f.websites.pc %>% 
  dplyr::select(-cumfreq) %>% 
  rename(Postcodes = level,
         F = freq,
         'F (%)' = perc,
         'Cummulative F' = cumperc)

@@

<!-- Multi-scale -->
To empirically test the two hypotheses stated in @sec-litreview, τhe website 
density metric was calculated for two different scales the Local Authority
Districts (LAD) and the Output Areas (OA). The former is an administrative unit 
and there are 374 LAD in the UK. The latter is the smallest census-based 
geographical unit and there are c. 230,000 of them in the UK. To create the 
website density metric, the yearly website counts at the LAD level are 
standardised by the number of firms in LAD to avoid biases associated with LAD 
hosting a large number of firms. Given that there is no such readily available 
statistic for the OA, the actual OA level counts are used. However, because of 
the the consistent spatial definition of OA (they host 40-250 households),[^3] 
website counts in OA are interpreted as a density metric too.

@@

>>>>

---
title: "A multi-scale story of the diffusion of a new technology: the web"
abstract: |
    This paper investigates the spatial diffusion of a new technology that is the Web in the UK. It employs novel data and machine learning methods to model the influence of well-established diffusion mechanisms. Contrary to previous studies, it adopts multiple scales, high spatial granularity and a long study period that captures the early stages of the Web until its maturity (1996-2012). Findings reveal the importance of such spatial mechanisms (namely distance, urban hierarchy and the S-shaped pattern of the cumulative level of adoption) even at granular scales. They also highlight spatial heterogeneity and instances of leapfrogging. 
keywords: 
  - diffusion
  - technologies
  - Web
  - digital
date: last-modified
bibliography: bibliography
format:
  elsevier-pdf:
    keep-tex: true
    text: |
      \usepackage[demo]{graphicx} % "demo" option just for this example
      \usepackage{subcaption}
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: authoryear
editor: 
  markdown: 
    wrap: 72
---

@@

<!-- geographers lack of interest for diffusion + gap -->
Economic geographers were always interested in how new technologies and
innovations diffuse across space and time with the work of
@hagerstrand1968innovation, which demonstrated the importance of
communications as a mechanism for the diffusion of innovations, being
the most impactful. It was followed by various empirical studies -- see
for instance @ormrod1990 and @iso2005. However, the torch of exploring
and modelling such diffusion processes had been passed to other
disciplines such as economics, business studies and sociology well
before the 'cultural turn' of economic geography
[@perkins2005international; @ding2010modeling].
<!-- NEW spatial granularity --> As a consequence, relevant studies
did not investigate diffusion mechanisms at granular geographical scales to 
provide local policy insights, for instance, regarding infrastructural 
development and local technology policies. 
A potential explanation of the lack of geographical studies exploring the 
diffusion of new and, more specifically for this paper, digital technologies 
across both space and time can be attributed to the scarcity of relevant and 
granular enough data, a problem also highlighted by @iso2005 and
@kemeny2011international. As @zook2022mapping pointed, digital
activities are hardly ever captured in official data.

<!-- contribution -->
This paper offers such a contribution: an economic geography study
revealing the importance of spatial mechanisms for the diffusion of a
digital technology that is the Web in the UK at a high level
of spatial granularity 
<!-- NEW TO JUSTIFY DATA TIME PERIOD-->
from the very early days until its maturity (1996-2012).
<!-- NEW DATA CLARIFICATION -->
It specifically focuses on the diffusion of commercial websites.
It does so by employing a novel source of big data and machine learning
algorithms to model the *active* engagement with Web -- that is creating and 
maintaining instead of just browsing a website -- during that period
and tests how different spatial and temporal mechanisms shaped its
diffusion. This paper also exemplifies how the combination of data
sources and state-of-the-art research methods which escape the
traditional social science domain can offer new lenses to geographical
research regarding the understanding of technological diffusion.

<!-- motivation: stakeholders -->
The motivation for this paper lies in the fact that there are various
stakeholders who are interested in knowing how new digital technologies
diffuse over space and time and use this knowledge to make predictions
about the diffusion of related future technologies. As per
@leibowicz2016representing, historical studies agree that technologies
diffuse differently in terms of times, rates, and geographies and can be
driven by related policies [@victor1993]. @meade2021modelling highlight
that a variety of actors have a direct interest in gaining such
knowledge including network equipment suppliers, network operators,
regulatory and local authorities. These processes and their effects vary
a lot across scales: although the diffusion of a new technology might
not be optimal at a local level, it might be beneficial from a global
perspective as it could lead to faster diffusion to less advantaged
places [@leibowicz2016representing]. Despite the spatial heterogeneity
of such diffusion mechanisms and the policy relevance, there is a
scarcity of studies analysing the diffusion of new technologies
at a detailed geographical level because of the lack of detailed enough
data and adequate research methods.

@@

<!-- NEW motivation: intangible/digital -->
Another line of motivation has to do with the very nature of the Web as a 
digital technology and how it evolved over time. The digital geographies field 
has debunked premature claims about the Internet's placeless-ness by mapping 
various of its facets and technological layers [@tranos2013death], but the lack 
of relevant long time-series data prevented researchers from observing digital
phenomena overtime. Most of these studies focused on its hard infrastructure 
[e.g. @malecki2002economic; @moss2000internet], on Internet subscribers 
[e.g. @blank2018local], or on social media users [e.g. @crampton2013beyond].
Exceptions include early work from @zook2000web, @zook2001old and 
@moss1997tracking, which utilised domain name registration data and use it to
map and illustrate the spatial footprint of the web content. Pivotal as these 
contributions have been in understanding the economic geographies of the 
Internet, they were based on cross-sectional, often spatially detailed, mapping 
and on short temporal comparisons (e.g. 1993-1998) and did not observe the 
evolution and diffusion of the Web until its maturity. As we know from the 
relevant literature, different technologies diffuse at different rates and 
there is no theory that can explain these differences 
[@leibowicz2016representing]. Hence, mapping and modelling the diffusion of the
Web can offer useful insights.

@@

<!-- description of the methods and results -->
To analyse these data, a diverse set of methods are employed. Firstly,
cumulative adoption curves are estimated for the UK as a whole and for
the Local Authority Districts (LAD), illustrating, for the first time,
such S-shaped diffusion patterns for local areas. The exploratory spatial data
analysis offers some first indications of the importance of the different 
spatial diffusion mechanisms in shaping the diffusion of the Web in the UK.
Importantly, the analysis adopts two distinct geographical scales: the LAD and 
the most detailed scale of UK census areas, the Output Areas (OA). Then, a
novel methodological framework based on Machine Learning (ML) algorithms
is developed to test the role of such mechanisms at these two scales. The 
results reveal, for the first time, the importance of different spatial 
diffusion mechanisms in the diffusion of a digital technology at very granular
geographical scales. The analysis also reveals that the diffusion of the Web 
is characterised by spatial heterogeneity, well-established spatial patterns 
and both stability and volatility.

<!-- contents -->
The structure of the paper goes as follows. @sec-litreview reviews the
mechanisms that shape the diffusion of a new technology and sets up the two 
hypotheses that are empirically tested in @sec-results. @sec-datamethods 
describes the data and methods used in this paper. The discussion of the 
results and the conclusions of the paper are offered in @sec-conclusions.

@@

<!-- hierarchy -->
@hagerstrand1968innovation also emphasised the role of the top-down
hierarchy of urban systems and highlighted that new technologies and
innovations are firstly adopted in larger cities and then diffuse to
second tier ones. This is a sequential instead of a simultaneous
process, which resembles the 'lead-lag' spatial acceleration effect in
market research [@bento2018time; @PERES201091]. Schmidt's Law
empirically illustrates a similar pattern [@grubler1990rise]. Core and
highly agglomerated regions is where new technologies are invented and
commercially deployed. This is where the first adopters tend to be
based. Then, technologies spread to the rim and eventually to the
periphery. Although adoption pace might be higher when new technologies
finally arrive to the periphery, the saturation levels may never
reach the ones in the core because of the lack of infrastructure or
other necessary institutions [@leibowicz2016representing].

@@

> H~2~: The diffusion of the commercial Web is shaped by both early adoption 
and latecomer effects.

These two hypotheses are empirically tested in @sec-results employing 
the data and methods described in the next section.

@@

[^2]: [https://archive.org/](https://archive.org/).

@@

The analysis is firstly conducted on websites, which only contain one
unique postcode. As a robustness check, the analysis is replicated for
an extended subset of websites, which include up to 10 unique postcodes
to capture commercial websites which point to multiple locations. They
are geolocated by equally attaching them to these locations. This
extended sample includes 94 percent of all the archived websites in
2000 (see Table \ref{f2000}).

@@

f.websites.pc <- f.websites.pc %>% 
  dplyr::select(-cumfreq) %>% 
  rename(Postcodes = level,
         F = freq,
         'F (%)' = perc,
         'cumulative F' = cumperc)

@@

<!-- Multi-scale -->
To empirically test the two hypotheses stated in @sec-litreview, the website 
density metric was calculated for two different scales the Local Authority
Districts (LAD) and the Output Areas (OA). The former is the lowest
administrative unit in the UK and there are 374 LAD. The latter is the smallest
census-based geographical unit and there are c. 230,000 of them in the UK. To 
create the website density metric, the yearly website counts at the LAD level 
are standardised by the number of firms in LAD to avoid biases associated with 
LAD hosting a large number of firms. Given that there is no such readily 
available statistic for the OA, the actual OA level counts are used. However, 
because of the the consistent spatial definition of OA (they host 40-250
households),[^3] website counts in OA are interpreted as a density metric too.

@@

<!-- CV -->
It needs to be highlighted here that the cross validation (CV) for all
models is spatially and temporally sensitive. Instead of using 10 random
and space- and time-agnostic folds, the `CAST` package is employed which
allows holding back data points from specific years and spatial units (regions)
and use them for testing in order to estimate the model performance
[@meyer2018improving].

<!-- assessment -->
To assess the predictive power of the model, three broadly utilised
metrics are employed: the coefficient of determination ($R^2$), mean
absolute error (MAE) and root mean square error (RMSE):
